Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-12-03 manuscript/manuscript.Rmd
tests more spp for QBM

Abstract
--------
Rapid climate change has generated growing interest in forecasts of future population trajectories. Traditional population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. An alternative is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. Here, we test the consequences of aggregating individual responses when forecasting climate change impacts on four perennial grass species in a semi-arid grassland in Montana, USA. We parameterized two population models based on the same dataset, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their accuracy, precision, and sensitivity to climate variables. The individual-level model was more accurate and precise than the aggregated model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model missed important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. Our analysis indicates that there is no shortcut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data efficiently across environmental gradients in space and time.

@@

### Comparison of forecast models
The IPM had lower overall error (MAE, mean absolute error) compared to the QBM for three species (Table 1). The IPM MAE is significantly lower at $\alpha=0.05$ for *B. gracilis* (*P* = `r mae_ps[1,"value"]`), *H. comata* (*P* = `r mae_ps[2,"value"]`), and *P. smithii* (*P* = `r mae_ps[3,"value"]`). MAEs are statisticially similar between models for *P. secunda* (*P* = `r mae_ps[4,"value"]`). *P* values are highly sensitive to sample size, so not entirely appropriate in simulation exercises where we control the samples size. But, for our purposes they serve as relatively unbiased comparison metrics. In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (1) lower MAE for three of the four species, (2) statistically similar MAE for the one other species, and (3) considerably more precise forecasts for all species.

@@

Tables
------
```{r table_1,results='asis',message=FALSE, comment=NA, echo=FALSE}
####
##  Make a table for average residual from each model and year effects structure
####

@@

Figures
-------
```{r figure_1, dependson="plot-options", fig.cap="Time series of average percent cover over all quadrats for our four focal species: *Bouteloua gracilis* (BOGR), *Hesperostipa comata* (HECO), *Pascopyrum smithii* (PASM), and *Poa secunda* (POSE). Light grey lines show trajectories of individual quadrats. Note the different y-axis scales across panels.", fig.height=3, fig.width=8.5}
obs_data <- read.csv("../analysis/speciesData/quadAllCover.csv")
#remove zeros for averaging
obs_data <- obs_data[which(obs_data$propCover>0),]
avgobs <- ddply(obs_data, .(year, Species), summarise,
                mean_cover = mean(propCover))

@@

```{r figure_2,dependson="plot-options", fig.cap="Work flow of the data aggregation, model fitting, and population simulating.", fig.height=6}
library(png)
library(grid)
img <- readPNG("components/figure/micromeso_flowchart.png")
grid.raster(img)
```

```{r figure_3, dependson="plot-options", fig.cap="Sensitivity of equilibrium cover simulated from the IPM to each climate scenario applied to individual and combined vital rates. For example, the points associated with G show the median cover from IPM simulations where a climate perturbation is applied only to the growth regression climate covariates. These simulations use mean parameter values for clarity.", fig.width=8, fig.height=6, cache=FALSE, results='hide'}
setwd("~/Repos/MicroMesoForecast/analysis/")

@@

```{r figure_4, dependson="plot-options", fig.cap="Effect of quadrat sample size on the precision (standard deviation) of main climate effect estimates in the QBM. Increasing the number of quadrats results in diminishing returns in terms of parameter certainty. Light dashed lines show individual climate effects at five quadrat sample sizes. Thick dark lines are inverse gaussian fits showing the mean effect of increasing quadrat sample size on parameter precision.", fig.width=3, fig.height=7.5, cache=FALSE, results='hide'}
library(mgcv)

@@

```{r figure_5, dependson="plot-options", fig.cap="Correlations (r) between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species. Across top panels, 'growth' = growth regression, 'rec' = recruitment regression, 'surv' = survival regression.", fig.width=6, fig.height=6, cache=FALSE, results='hide'}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

@@

```{r figure_6, dependson="plot-options", fig.cap="Mean (points) and 90% quantiles (errorbars) for the proportional difference between baseline simulations (using observed climate) and the climate pertubation simulation on the x-axis. We calculated proportional difference as log(perturbed climate cover) - log(observed climate cover), where 'perturbed' and 'observed' refer to the climate time series used to drive interannual variation in the simulations. Model error and parameter uncertainty were propagated through the simulation phase. Climate simulations are as in Figure 3.", fig.width=4.5, fig.height=7, cache=FALSE, results='hide'}
####
##  Vary parameter runs
####
setwd("~/Repos/MicroMesoForecast/analysis/")
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
##  IPM 
path <- "./ipm/simulations/results/climchange_varyparams/"
ipmfiles_mean <- list.files(path)
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
torm <- c(grep("10", ipmfiles_mean), grep("20", ipmfiles_mean), grep("30", ipmfiles_mean))
ipmfiles_mean <- ipmfiles_mean[-torm]
for(file in ipmfiles_mean){
  tmp <- read.csv(paste(path,file,sep=""))
  tmp$model <- "IPM"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]

>>>>

Summary
-------
1. Rapid climate change has generated growing interest in forecasts of future population trajectories. Traditional population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. An alternative is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. 
2. We test the consequences of aggregating individual responses when forecasting climate change impacts on two perennial grass species (*Pascopyrum smithii* and *Sprobolis flexuosus*). We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their accuracy, precision, and sensitivity to climate variables. For both models we used Bayesian ridge regression to identify the optimal predictive model in terms of climate covariate strengths. 
3. The individual-level model was more accurate and precise than the aggregated model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model missed important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. 
4. Our analysis indicates that there is no shortcut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data efficiently across environmental gradients in space and time.

@@

### Comparison of forecast models

@@

Tables
------
```{r table_1,results='asis',message=FALSE, comment=NA, echo=FALSE, eval=F}
####
##  Make a table for average residual from each model and year effects structure
####

@@

Figures
-------
```{r figure_1, dependson="plot-options", fig.cap="Time series of average percent cover over all quadrats for our four focal species: *Bouteloua gracilis* (BOGR), *Hesperostipa comata* (HECO), *Pascopyrum smithii* (PASM), and *Poa secunda* (POSE). Light grey lines show trajectories of individual quadrats. Note the different y-axis scales across panels.", fig.height=3, fig.width=8.5, eval=F}
obs_data <- read.csv("../analysis/speciesData/quadAllCover.csv")
#remove zeros for averaging
obs_data <- obs_data[which(obs_data$propCover>0),]
avgobs <- ddply(obs_data, .(year, Species), summarise,
                mean_cover = mean(propCover))

@@

```{r figure_2,dependson="plot-options", fig.cap="Work flow of the data aggregation, model fitting, and population simulating.", fig.height=6, eval=F}
library(png)
library(grid)
img <- readPNG("components/figure/micromeso_flowchart.png")
grid.raster(img)
```

```{r figure_3, dependson="plot-options", fig.cap="Sensitivity of equilibrium cover simulated from the IPM to each climate scenario applied to individual and combined vital rates. For example, the points associated with G show the median cover from IPM simulations where a climate perturbation is applied only to the growth regression climate covariates. These simulations use mean parameter values for clarity.", fig.width=8, fig.height=6, cache=FALSE, results='hide', eval=F}
setwd("~/Repos/MicroMesoForecast/analysis/")

@@

```{r figure_4, dependson="plot-options", fig.cap="Effect of quadrat sample size on the precision (standard deviation) of main climate effect estimates in the QBM. Increasing the number of quadrats results in diminishing returns in terms of parameter certainty. Light dashed lines show individual climate effects at five quadrat sample sizes. Thick dark lines are inverse gaussian fits showing the mean effect of increasing quadrat sample size on parameter precision.", fig.width=3, fig.height=7.5, cache=FALSE, results='hide', eval=F}
library(mgcv)

@@

```{r figure_5, dependson="plot-options", fig.cap="Correlations (r) between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species. Across top panels, 'growth' = growth regression, 'rec' = recruitment regression, 'surv' = survival regression.", fig.width=6, fig.height=6, cache=FALSE, results='hide', eval=F}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

@@

```{r figure_6, dependson="plot-options", fig.cap="Mean (points) and 90% quantiles (errorbars) for the proportional difference between baseline simulations (using observed climate) and the climate pertubation simulation on the x-axis. We calculated proportional difference as log(perturbed climate cover) - log(observed climate cover), where 'perturbed' and 'observed' refer to the climate time series used to drive interannual variation in the simulations. Model error and parameter uncertainty were propagated through the simulation phase. Climate simulations are as in Figure 3.", fig.width=4.5, fig.height=7, cache=FALSE, results='hide', eval=F}
####
##  Vary parameter runs
####
setwd("~/Repos/MicroMesoForecast/analysis/")
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
##  IPM 
path <- "./ipm/simulations/results/climchange_varyparams/"
ipmfiles_mean <- list.files(path)
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
torm <- c(grep("10", ipmfiles_mean), grep("20", ipmfiles_mean), grep("30", ipmfiles_mean))
ipmfiles_mean <- ipmfiles_mean[-torm]
for(file in ipmfiles_mean){
  tmp <- read.csv(paste(path,file,sep=""))
  tmp$model <- "IPM"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]

### 2015-12-03 manuscript/manuscript.tex
tests more spp for QBM

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

%%% Change title format to be more compact
\usepackage{titling}
\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

\subsection{Abstract}\label{abstract}

Rapid climate change has generated growing interest in forecasts of
future population trajectories. Traditional population models, typically
built using detailed demographic observations from one study site, can
address climate change impacts at one location, but are difficult to
scale up to the landscape and regional scales relevant to management
decisions. An alternative is to build models using population-level data
that is much easier to collect over broad spatial scales than
individual-level data. However, such models ignore the fact that climate
drives population growth through its influence on individual
performance. Here, we test the consequences of aggregating individual
responses when forecasting climate change impacts on four perennial
grass species in a semi-arid grassland in Montana, USA. We parameterized
two population models based on the same dataset, one based on
individual-level data (survival, growth and recruitment) and one on
population-level data (percent cover), and compared their accuracy,
precision, and sensitivity to climate variables. The individual-level
model was more accurate and precise than the aggregated model when
predicting out-of-sample observations. When comparing climate effects
from both models, the population-level model missed important climate
effects from at least one vital rate for each species. Increasing the
sample size at the population-level would not necessarily reduce
forecast uncertainty; the way to reduce uncertainty is to capture unique
climate dependence of individual vital rates. Our analysis indicates
that there is no shortcut to forecasting climate change impacts on plant
populations --- detailed demographic data is essential. Despite the
superiority of the individual-level model, the forecasts it generated
still were too uncertain to be useful for decision-makers. We need new
methods to collect demographic data efficiently across environmental
gradients in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). The increasing
statistical sophistication of population models (Rees and Ellner 2009)
makes them promising tools for predicting the impacts of environmental
change on species persistence and abundance. But reconciling the scales
at which population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The problem is
that most population models are built using data from a single study
site because collecting those data, which involves tracking the fates of
individuals plants, is so difficult. The resulting models cannot be
applied to the landscape and regional scales relevant to decision-making
without information about how the fitted parameters respond to spatial
variation in biotic and abiotic drivers (Sæther et al. 2007). The
limited spatial extent of individual-level demographic datasets
constrains our ability to use population models to address applied
questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions. However, the limitations of SDMs are well known
(Pearson and Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson
2012). Ideally, researchers would provide managers with landscape-scale
population models, combining the extent of SDMs with information about
dynamics and species abundances (Schurr et al. 2012, Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bjørnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially since such
data are feasible to collect across broad spatial extents (e.g.,
Queenborough et al. 2011). The choice between individual and
population-level data involves a difficult trade-off: while
individual-level data leads to more mechanistic models, population-level
data leads to models that can be applied over greater spatial and
temporal extents. An open question is how much forecasting skill is lost
when we build models based on population rather than individual-level
data.

@@

The IPM had lower overall error (MAE, mean absolute error) compared to
the QBM for three species (Table 1). The IPM MAE is significantly lower
at \(\alpha=0.05\) for \emph{B. gracilis} (\emph{P} = 0.0012), \emph{H.
comata} (\emph{P} = 4.0586 × 10-8), and \emph{P. smithii} (\emph{P} =
3.183 × 10-5). MAEs are statisticially similar between models for
\emph{P. secunda} (\emph{P} = 0.0922). \emph{P} values are highly
sensitive to sample size, so not entirely appropriate in simulation
exercises where we control the samples size. But, for our purposes they
serve as relatively unbiased comparison metrics. In no case did the QBM
significantly outperform the IPM (Table 1). The IPM was consistently
more precise than the QBM, with lower distances between the 90\%
quantiles across all species (Table 1). In general the IPM outperformed
the QBM because it had (1) lower MAE for three of the four species, (2)
statistically similar MAE for the one other species, and (3)
considerably more precise forecasts for all species.

@@

Our interpretation is that the QBM is ``missing'' climate signals
associated with at least one vital rate for each species. This leads to
inaccurate and imprecise forecasts because the QBM statistical model
struggles to explain variation due to climate variables that have
positive and negative impacts on different vital rates. When this is the
case, as it is for all our species to varying degrees (Fig. 3),
forecasts from models based on population-level data will fail. Our
result is consistent with related work on the importance of
individual-level data to forecast population responses to exogenous
drivers (Clark et al. 2011a, 2011b, 2012, Galván et al. 2014).

@@

\begin{table}[ht]
\centering
\caption{Accuracy (mean absolute error, MAE) and precision (90\% Distance) of out of sample predictions. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90\% Distance refers to the average distance between the upper and lower 90th percentiles of the 100 predicted values for each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 12.18 & 38.52 & 9.43 \\ 
  BOGR & QBM & 19.66 & 56.50 & 9.26 \\ 
  HECO & IPM & 1.22 & 6.47 & 1.15 \\ 
  HECO & QBM & 12.35 & 41.11 & 1.18 \\ 
  PASM & IPM & 0.19 & 1.65 & 0.42 \\ 
  PASM & QBM & 0.55 & 7.78 & 0.42 \\ 
  POSE & IPM & 1.37 & 7.64 & 1.25 \\ 
  POSE & QBM & 1.79 & 40.59 & 1.27 \\ 
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Time series of average percent cover over all quadrats for our
four focal species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa
comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda}
(POSE). Light grey lines show trajectories of individual quadrats. Note
the different y-axis scales across panels.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Work flow of the data aggregation, model fitting, and
population simulating.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{Sensitivity of equilibrium cover simulated from the IPM to each
climate scenario applied to individual and combined vital rates. For
example, the points associated with G show the median cover from IPM
simulations where a climate perturbation is applied only to the growth
regression climate covariates. These simulations use mean parameter
values for clarity.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4.pdf}
\caption{Effect of quadrat sample size on the precision (standard
deviation) of main climate effect estimates in the QBM. Increasing the
number of quadrats results in diminishing returns in terms of parameter
certainty. Light dashed lines show individual climate effects at five
quadrat sample sizes. Thick dark lines are inverse gaussian fits showing
the mean effect of increasing quadrat sample size on parameter
precision.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_5.pdf}
\caption{Correlations (r) between QBM and IPM estimates of climate
effects. We ignore sizeXclimate interactions since these are not
directly comparable across model types. The QBM does not have multiple
vital rates, so its values are repeated across panels within each
species. Across top panels, `growth' = growth regression, `rec' =
recruitment regression, `surv' = survival regression.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Mean (points) and 90\% quantiles (errorbars) for the
proportional difference between baseline simulations (using observed
climate) and the climate pertubation simulation on the x-axis. We
calculated proportional difference as log(perturbed climate cover) -
log(observed climate cover), where `perturbed' and `observed' refer to
the climate time series used to drive interannual variation in the
simulations. Model error and parameter uncertainty were propagated
through the simulation phase. Climate simulations are as in Figure 3.}
\end{figure}

@@

Araújo, M. B., and A. T. Peterson. 2012. Uses and misuses of bioclimatic
envelope modeling. Ecology 93:1527--1539.

@@

Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
recruitment stage to stabilize grassland coexistence. Ecological
Monographs.

@@

Clark, J. S., and O. N. Bjørnstad. 2004. Population time series: Process
variability, observation errors, missing values, lags, and hidden
states. Ecology 85:3140--3150.

@@

Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
HilleRisLambers, I. Ibáñez, S. LaDeau, S. McMahon, J. Metcalf, J. Mohan,
E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and P.
Wyckoff. 2010. High-dimensional coexistence based on individual
variation: a synthesis of evidence. Ecological Monographs 80:569--608.

@@

Galván, J. D., J. J. Camarero, and E. Gutiérrez. 2014. Seeing the trees
for the forest: Drivers of individual growth responses to climate in
Pinus uncinata mountain forests. Journal of Ecology 102:1244--1257.

@@

Petchey, O. L., M. Pontarp, T. M. Massie, S. Kéfi, A. Ozgul, M.
Weilenmann, G. M. Palamara, F. Altermatt, B. Matthews, J. M. Levine, D.
Z. Childs, B. J. McGill, M. E. Schaepman, B. Schmid, P. Spaak, A. P.
Beckerman, F. Pennekamp, and I. S. Pearse. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova, R. B.
O'Hara, F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schröder, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.

@@

Sæther, B. E., S. Engen, V. Grøtan, W. Fiedler, E. Matthysen, M. E.
Visser, J. Wright, A. P. Møller, F. Adriaensen, H. {Van Balen}, D.
Balmer, M. C. Mainwaring, R. H. McCleery, M. Pampus, and W. Winkel.
2007. The extended Moran effect and large-scale synchronous fluctuations
in the size of great tit and blue tit populations. Journal of Animal
Ecology 76:315--325.

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

\subsection{Summary}\label{summary}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Rapid climate change has generated growing interest in forecasts of
  future population trajectories. Traditional population models,
  typically built using detailed demographic observations from one study
  site, can address climate change impacts at one location, but are
  difficult to scale up to the landscape and regional scales relevant to
  management decisions. An alternative is to build models using
  population-level data that is much easier to collect over broad
  spatial scales than individual-level data. However, such models ignore
  the fact that climate drives population growth through its influence
  on individual performance.
\item
  We test the consequences of aggregating individual responses when
  forecasting climate change impacts on two perennial grass species
  (\emph{Pascopyrum smithii} and \emph{Sprobolis flexuosus}). We
  parameterized two population models for each species, one based on
  individual-level data (survival, growth and recruitment) and one on
  population-level data (percent cover), and compared their accuracy,
  precision, and sensitivity to climate variables. For both models we
  used Bayesian ridge regression to identify the optimal predictive
  model in terms of climate covariate strengths.
\item
  The individual-level model was more accurate and precise than the
  aggregated model when predicting out-of-sample observations. When
  comparing climate effects from both models, the population-level model
  missed important climate effects from at least one vital rate for each
  species. Increasing the sample size at the population-level would not
  necessarily reduce forecast uncertainty; the way to reduce uncertainty
  is to capture unique climate dependence of individual vital rates.
\item
  Our analysis indicates that there is no shortcut to forecasting
  climate change impacts on plant populations --- detailed demographic
  data is essential. Despite the superiority of the individual-level
  model, the forecasts it generated still were too uncertain to be
  useful for decision-makers. We need new methods to collect demographic
  data efficiently across environmental gradients in space and time.
\end{enumerate}

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). The increasing
statistical sophistication of population models (Rees and Ellner 2009)
makes them promising tools for predicting the impacts of environmental
change on species persistence and abundance. But reconciling the scales
at which population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The problem is
that most population models are built using data from a single study
site because collecting those data, which involves tracking the fates of
individuals plants, is so difficult. The resulting models cannot be
applied to the landscape and regional scales relevant to decision-making
without information about how the fitted parameters respond to spatial
variation in biotic and abiotic drivers (S{æ}ther et al. 2007). The
limited spatial extent of individual-level demographic datasets
constrains our ability to use population models to address applied
questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions. However, the limitations of SDMs are well known
(Pearson and Dawson 2003, Elith and Leathwick 2009, Ara{ú}jo and
Peterson 2012). Ideally, researchers would provide managers with
landscape-scale population models, combining the extent of SDMs with
information about dynamics and species abundances (Schurr et al. 2012,
Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bj{ø}rnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially since such
data are feasible to collect across broad spatial extents (e.g.,
Queenborough et al. 2011). The choice between individual and
population-level data involves a difficult trade-off: while
individual-level data leads to more mechanistic models, population-level
data leads to models that can be applied over greater spatial and
temporal extents. An open question is how much forecasting skill is lost
when we build models based on population rather than individual-level
data.

@@

@@

Our interpretation is that the QBM is ``missing'' climate signals
associated with at least one vital rate for each species. This leads to
inaccurate and imprecise forecasts because the QBM statistical model
struggles to explain variation due to climate variables that have
positive and negative impacts on different vital rates. When this is the
case, as it is for all our species to varying degrees (Fig. 3),
forecasts from models based on population-level data will fail. Our
result is consistent with related work on the importance of
individual-level data to forecast population responses to exogenous
drivers (Clark et al. 2011a, 2011b, 2012, Galv{á}n et al. 2014).

@@

@@

@@

Ara{ú}jo, M. B., and A. T. Peterson. 2012. Uses and misuses of
bioclimatic envelope modeling. Ecology 93:1527--1539.

@@

Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
recruitment stage to stabilize grassland coexistence. Ecological
Monographs 85:373--392.

@@

Clark, J. S., and O. N. Bj{ø}rnstad. 2004. Population time series:
Process variability, observation errors, missing values, lags, and
hidden states. Ecology 85:3140--3150.

@@

Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
HilleRisLambers, I. Ib{á}{ñ}ez, S. LaDeau, S. McMahon, J. Metcalf, J.
Mohan, E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and
P. Wyckoff. 2010. High-dimensional coexistence based on individual
variation: a synthesis of evidence. Ecological Monographs 80:569--608.

@@

Galv{á}n, J. D., J. J. Camarero, and E. Guti{é}rrez. 2014. Seeing the
trees for the forest: Drivers of individual growth responses to climate
in Pinus uncinata mountain forests. Journal of Ecology 102:1244--1257.

@@

Petchey, O. L., M. Pontarp, T. M. Massie, S. K{é}fi, A. Ozgul, M.
Weilenmann, G. M. Palamara, F. Altermatt, B. Matthews, J. M. Levine, D.
Z. Childs, B. J. McGill, M. E. Schaepman, B. Schmid, P. Spaak, A. P.
Beckerman, F. Pennekamp, and I. S. Pearse. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova, R. B.
O'Hara, F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schr{ö}der, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.

@@

S{æ}ther, B. E., S. Engen, V. Gr{ø}tan, W. Fiedler, E. Matthysen, M. E.
Visser, J. Wright, A. P. M{ø}ller, F. Adriaensen, H. {Van Balen}, D.
Balmer, M. C. Mainwaring, R. H. McCleery, M. Pampus, and W. Winkel.
2007. The extended Moran effect and large-scale synchronous fluctuations
in the size of great tit and blue tit populations. Journal of Animal
Ecology 76:315--325.
